---
title: "Creating and Sharing Recipes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating and Sharing Recipes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Why Recipes?

Anyone who works with household survey microdata knows this pattern: you download the raw files, open the codebook, and spend days recoding employment status, harmonizing income variables, and building indicators. Months later, a colleague starts the same project and writes the same code from scratch.

In STATA, teams share `.do` files, but these are tightly coupled to specific file paths and variable names, and there is no standard way to discover or validate them.

**Recipes** are metasurvey's answer to this problem. A recipe is a portable, documented, and validated collection of transformation steps that can:

- Be applied to any compatible edition of a survey with a single function call
- Be published to a registry where others can discover and reuse them
- Be certified by institutions for official use
- Generate automatic documentation of inputs, outputs, and pipeline

## Recipe Lifecycle

```
  ┌──────────────┐     ┌──────────────┐     ┌──────────────┐
  │  1. Develop   │────▶│  2. Package   │────▶│  3. Validate  │
  │  steps on a   │     │  steps into   │     │  against new  │
  │  survey       │     │  a recipe     │     │  data         │
  └──────────────┘     └──────────────┘     └──────┬───────┘
                                                    │
  ┌──────────────┐     ┌──────────────┐             │
  │  5. Discover  │◀────│  4. Publish   │◀────────────┘
  │  & reuse      │     │  to registry  │
  │  recipes      │     │  or API       │
  └──────────────┘     └──────────────┘
```

## Loading a Survey

The typical starting point is loading a survey with `load_survey()`. You can load example data or point to your own files:

```{r load-survey, eval = FALSE}
library(metasurvey)

# Load ECH 2022 with example data
ech_2022 <- load_survey(
  load_survey_example("ech", "ech_2022"),
  svy_type    = "ech",
  svy_edition = "2022",
  svy_weight  = add_weight(annual = "pesoano")
)

# Or load with existing recipes from the registry
ech_2022 <- load_survey(
  load_survey_example("ech", "ech_2022"),
  svy_type    = "ech",
  svy_edition = "2022",
  svy_weight  = add_weight(annual = "pesoano"),
  recipes     = get_recipe("ech", "2022")
)
```

## Building a Recipe from Steps

The most common workflow consists of developing transformations interactively on a survey and then converting the recorded steps into a recipe.

```{r build-recipe}
library(metasurvey)
library(data.table)

set.seed(42)
n <- 200

# Simulate survey microdata (standing in for load_survey)
dt <- data.table(
  id = 1:n,
  age = sample(18:80, n, replace = TRUE),
  sex = sample(c(1, 2), n, replace = TRUE),
  income = round(runif(n, 5000, 80000)),
  activity = sample(c(2, 3, 5, 6), n,
    replace = TRUE,
    prob = c(0.55, 0.05, 0.05, 0.35)
  ),
  weight = round(runif(n, 0.5, 3.0), 4)
)

svy <- Survey$new(
  data    = dt,
  edition = "2023",
  type    = "ech",
  psu     = NULL,
  engine  = "data.table",
  weight  = add_weight(annual = "weight")
)

# Develop transformations interactively
svy <- step_compute(svy,
  income_thousands = income / 1000,
  employed = ifelse(activity == 2, 1L, 0L),
  comment = "Income scaling and employment indicator"
)

svy <- step_recode(svy, labor_status,
  activity == 2 ~ "Employed",
  activity %in% 3:5 ~ "Unemployed",
  activity %in% 6:8 ~ "Inactive",
  .default = "Other",
  comment = "ILO labor force classification"
)

svy <- step_recode(svy, age_group,
  age < 25 ~ "Youth",
  age < 45 ~ "Adult",
  age < 65 ~ "Mature",
  .default = "Senior",
  comment = "Standard age groups"
)

# Convert all steps to a recipe
labor_recipe <- steps_to_recipe(
  name        = "Labor Force Indicators",
  user        = "Research Team",
  svy         = svy,
  description = "Standard labor force indicators following ILO definitions",
  steps       = get_steps(svy),
  topic       = "labor"
)

labor_recipe
```

## Recipe Documentation

Every recipe can automatically generate its documentation from its steps. The `doc()` method returns a list with input variables, output variables, and the step-by-step pipeline:

```{r recipe-doc}
doc <- labor_recipe$doc()
names(doc)
```

```{r recipe-doc-detail}
# What variables does the recipe need?
doc$input_variables

# What variables does it create?
doc$output_variables

# Step-by-step pipeline
doc$pipeline
```

This documentation is generated automatically, with no manual effort required.

## Validation

Before applying a recipe to new data, verify that all required variables exist. The `validate()` method stops with a clear error if any dependency is missing:

```{r recipe-validate}
labor_recipe$validate(svy)
```

## Applying Recipes to a Survey

Attach one or more recipes to a survey and apply them with `bake_recipes()`:

```{r apply-recipe}
# Create a fresh survey with same structure (simulating a new edition)
set.seed(99)
dt2 <- data.table(
  id = 1:100,
  age = sample(18:80, 100, replace = TRUE),
  sex = sample(c(1, 2), 100, replace = TRUE),
  income = round(runif(100, 5000, 80000)),
  activity = sample(c(2, 3, 5, 6), 100,
    replace = TRUE,
    prob = c(0.55, 0.05, 0.05, 0.35)
  ),
  weight = round(runif(100, 0.5, 3.0), 4)
)

svy2 <- Survey$new(
  data = dt2, edition = "2024", type = "ech",
  psu = NULL, engine = "data.table",
  weight = add_weight(annual = "weight")
)

# Attach and bake
svy2 <- add_recipe(svy2, labor_recipe)
svy2 <- bake_recipes(svy2)

head(get_data(svy2)[, .(id, income_thousands, labor_status, age_group)], 5)
```

The same recipe applied to a different edition produces consistent results. This is how metasurvey ensures reproducibility over time.

In practice, you can load a survey and apply published recipes in a single call:

```{r load-with-recipes, eval = FALSE}
# Load ECH 2023 and apply the labor recipe from the registry
ech_2023 <- load_survey(
  load_survey_example("ech", "ech_2023"),
  svy_type    = "ech",
  svy_edition = "2023",
  svy_weight  = add_weight(annual = "pesoano"),
  recipes     = get_recipe("ech", "2023", topic = "labor_market"),
  bake        = TRUE
)
```

## Categories

Categories help organize recipes by topic:

```{r categories}
cats <- default_categories()
vapply(cats, function(c) c$name, character(1))
```

Add categories to a recipe using `add_category()`:

```{r add-category}
labor_recipe <- add_category(labor_recipe, "labor_market", "Labor market analysis")
labor_recipe <- add_category(labor_recipe, "income", "Income-related indicators")
labor_recipe
```

## Certification

The certification system offers three levels of trust:

| Level | Meaning |
|-------|---------|
| `community` | User contribution (default), no review |
| `reviewed` | Peer-reviewed by a recognized team |
| `official` | Endorsed for official statistics |

Higher certification levels appear first in search results and signal that the recipe has been reviewed.

## Publishing and Discovering Recipes

The real power of recipes lies in sharing them. Every recipe you create can be published to the **metasurvey registry**, where other researchers can discover, reuse, and build upon your work.

### Publishing to the Public Registry

The recommended workflow is to publish recipes to the public API. Anyone can browse recipes without an account; publishing requires registration:

```{r api-sharing, eval = FALSE}
# One-time: register and authenticate
api_register("Your Name", "you@example.com", "password")
api_login("you@example.com", "password")

# Publish your recipe (your profile is attached automatically)
api_publish_recipe(labor_recipe)
```

When authenticated, `api_publish_recipe()` attaches your user profile to the recipe. Other users see who published it, along with institutional affiliation and certification level. This builds accountability and trust in shared recipes.

### Browsing and Searching

No authentication is needed to browse and download recipes:

```{r api-browse, eval = FALSE}
# Browse all ECH recipes
ech_recipes <- api_list_recipes(survey_type = "ech")

# Get a specific recipe by ID
r <- api_get_recipe(id = "recipe_id_here")
```

The tidy API provides convenient search and filtering:

```{r tidy-api-demo, eval = FALSE}
# Search by text
found <- search_recipes("labor")

# Filter by survey type and certification
official <- filter_recipes(survey_type = "ech", certification_level = "official")

# Rank by popularity
top <- rank_recipes(n = 10)
```

### Interactive Explorer

The Shiny app provides a visual interface for browsing recipes and workflows:

```{r shiny-app, eval = FALSE}
run_metasurvey_app()
```

The explorer shows recipe cards with certification badges, download counts, and pipeline previews. Clicking a recipe opens a detail view with the full pipeline, an R code snippet, and links to related workflows.

### Private Registry for Institutions

Institutions that work with **confidential or restricted-access surveys** may need a private registry. metasurvey supports this via a self-hosted backend with MongoDB:

```{r private-registry, eval = FALSE}
# Point to your institution's private API
configure_api("https://your-institution.example.com/api")

# From here, the workflow is identical
api_login("analyst@institution.edu", "password")
api_publish_recipe(labor_recipe)
api_list_recipes(survey_type = "ech")
```

See the [API and Database](api-database.html) vignette for instructions on deploying the Plumber API with MongoDB for your own organization.

## Best Practices

1. **Name recipes descriptively** -- include the survey type and topic
   (e.g., `"ECH Labor Force Indicators"`).
2. **Add descriptions** -- document what the recipe computes and why.
3. **Use categories and topics** -- make recipes easier to discover.
4. **Validate before sharing** -- call `validate()` on sample data to
   ensure all dependencies exist.
5. **Version your recipes** -- use `set_version()` when updating them.

## Next Steps

- **[Estimation Workflows](workflows-and-estimation.html)** -- Use `workflow()` to compute weighted estimates from processed data
- **[ECH Case Study](ech-case-study.html)** -- See recipes in action in a real labor market analysis
- **[Getting Started](getting-started.html)** -- Review the basics of steps and Survey objects
